\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-08-11 tex/ms.tex
typo
better english

\section{Variational Inference}
\label{sec:variational}
%
\xxx{Explain what we're doing and the connection between our GP and VI.}

The mean $\pmb{\mu}$ and covariance $\pmb{\Sigma}$ of the flux GP are given by
%
\begin{align}
    \label{eq:mean}
    \pmb{\mu}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{f}(\pmb{\theta}) \, \Big| \, \pmb{\theta} \Big]
    \\
    \label{eq:cov}
    \pmb{\Sigma}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{f}(\pmb{\theta}) \, \mathbf{f}^\top(\pmb{\theta}) \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}^2(\pmb{\theta})
\end{align}
%
where $\mathrm{E}[\cdots]$ denotes the expectation value,
$\mathbf{f}$ is a random vector-valued variable corresponding to
a set of flux measurements, $\pmb{\theta}$ is a vector
of hyperparameters on which the GP depends, and the squaring operation in
Equation~(\ref{eq:cov}) is performed element-wise.
The vector $\pmb{\theta}$ includes physical properties of the star such
as its inclination $i$ and rotational period $P$ as well as parameters
describing the shape of the probability density function (PDF) governing
the distribution of features on the surface.
%
Since the flux
$\mathbf{f}$ is linear in
the spherical harmonic coefficient vector $\mathbf{y}$,
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f}(\pmb{\theta}) = \mathbf{A}(\pmb{\theta}) \, \mathbf{y}
    \quad,
\end{align}
%
where $\mathbf{A}$ is the \starry design matrix \citep{Luger2019},
we may write the mean and covariance as
%
\begin{align}
    \pmb{\mu}(\pmb{\theta})
     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
    \\
    \pmb{\Sigma}(\pmb{\theta})
     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\Sigma}_{\mathbf{y}} \, \mathbf{A}^\top(\pmb{\theta})
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:mean_y}
    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
    \\
    \label{eq:cov_y}
    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}_{\mathbf{y}}^2(\pmb{\theta})
\end{align}
%
are the mean and covariance of the spherical harmonic GP.
The expectation values in Equations~(\ref{eq:mean_y}) and (\ref{eq:cov_y})
above are given by the integrals
%
\begin{align}
    \label{eq:exp_y}
    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
    \\
    \label{eq:exp_yy}
    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
    \quad,
\end{align}
%
where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
distribution of features on the surface and
$p(\mathbf{x} \, \big| \, \pmb{\theta})$ is its PDF.

The goal of this paper is to show that for suitable choices of $\mathbf{y}(\mathbf{x})$

@@

    \mathbf{x} = \big\{ \varphi, \lambda, \delta, r \big\}

@@

    \mathbf{s}(\delta, r)

@@

where $\lambda$ is the longitude of a spot, $\phi$ is its latitude, $\delta$
is its contrast, and $r$ is its radius. The vector function $\mathbf{s}(\delta, r)$
returns the spherical harmonic expansion of a circular spot at $\lambda = \phi = 0$,

@@

$\mathbf{x}$ describing the size, contrast, and position of a spot to the

@@

    p(\delta \, \big| \, \pmb{\theta}_{\delta}) \,

@@

    \pmb{\theta} = \big\{
    \pmb{\theta}_{\lambda}, \,
    \pmb{\theta}_{\phi}, \,
    \pmb{\theta}_{\delta}, \,
    \pmb{\theta}_{r} \big\}

@@

    \mathbf{e_3}(\pmb{\theta})

@@

    \mathbf{E_3}(\pmb{\theta})

@@

    \mathbf{e_1}(\pmb{\theta}_r, \pmb{\theta}_\delta)
    \iint
    \mathbf{s}(\delta, r) \,

@@

    \mathrm{d}r \,
    p(\delta \, \big| \, \pmb{\theta}_{\delta}) \,
    \mathrm{d}\delta

@@

    \mathbf{e_2}(\pmb{\theta}_\phi, \mathbf{S})

@@

    \mathbf{e_3}(\pmb{\theta}_\lambda, \mathbf{P})

@@

@@

    \mathbf{E_1}(\pmb{\theta}_r, \pmb{\theta}_\delta)
    \iint
    \mathbf{s}(\delta, r) \, \mathbf{s}^\top(\delta, r) \,

@@

    \mathrm{d}r \,
    p(\delta \, \big| \, \pmb{\theta}_{\delta})
    \mathrm{d}\delta

@@

    \mathbf{E_2}(\pmb{\theta}_\phi, \mathbf{S})

@@

    \mathbf{E_3}(\pmb{\theta}_\lambda, \mathbf{P})

@@

@@

We devote the next three sections to the computation of these

@@

%
\section{The Spot Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
\label{sec:spot}
%
\subsection{Preliminaries}
%
\subsubsection{Functional form}
\label{sec:spot-function}
%
We adopt the following expression for the $n^{\mathrm{th}}$ term in $\mathbf{s}$
in the expansion of a spot at
$\theta = \varphi = 0$:
%
\begin{align}
    \label{eq:s}
    s_{n}(\delta, r) =
    \begin{cases}
        1 - \dfrac{\delta r'}{2 (1 + r')}
         & l = m = 0    \\[2em]
        -\dfrac{\delta r' \left( 2 + r' \right)}
        {2 \sqrt{2l + 1} (1 + r')^{l + 1}}
         & l > 0, m = 0 \\[2em]
        0
         & m \ne 0
    \end{cases}
\end{align}
%
where
%
\begin{align}
    \label{eq:rprime}
    r' \equiv c_0 + c_1 r
\end{align}
%
and $c_0, c_1 > 0$ are scaling constants (defined in more detail below).
In the expressions above,
$\delta \in (-\infty, 1]$ is the spot contrast (the fractional
decrease in the brightness
at the center of the spot) and $r \in [0, 1]$ is the normalized spot radius.
The expression in Equation~(\ref{eq:s}) is convenient because it satisfies
six important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity is azimuthally symmetric about the spot center
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $1 - \delta$
    \item The surface intensity at the antipode of the spot center is unity
    \item The size of the spot increases monotonically with $r$
\end{enumerate}
%
These properties may be demonstrated by considering the
expression for the surface intensity at a given point $(\theta, \varphi)$:
%
\begin{align}
    \begin{split}
        I(\theta, \varphi)
        & =
        \sum\limits_{n=0}^\infty
        s_{n} Y_{n}(\theta, \varphi) \\
        & =
        \sum\limits_{n=0}^\infty
        \begin{cases}
            s_{n} \sqrt{2l + 1} P_l(\cos\theta)
              & m = 0              \\[2em]
            0 & \mathrm{otherwise}
        \end{cases}
    \end{split}
\end{align}
%
where $Y_n$ is the spherical harmonic of degree $l(n)$ and order $m(n)$
(see Equation~\ref{eq:lm}),
$P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Note that the expression above is independent of the azimuth $\varphi$,
a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
expansion is therefore azimuthally symmetric about the spot center, as
stated in \textbf{1}.
%
Combining the above expression with Equation~(\ref{eq:s}) and
rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    I(\theta) =
    1 + \dfrac{\delta r'}{2}
    -
    \dfrac{\delta r' \left( 2 + r' \right)}{2 (1 + r')}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + r'}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
    %
    \\
    \intertext{where}
    %
    \begin{split}
        A & = 1 + \dfrac{\delta r'}{2}                      \\
        B & = \delta r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
        C & = \dfrac{1 + (1 + r')^2}{2 + 2 r'}
    \end{split}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{2}. The value at the minimum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
        & = 1 - \delta
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{3}, and the value at the maximum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
        & = 1
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{4}.
Finally, to show \textbf{5}, let us compute the half width at half minimum
$\Delta\theta$ of the intensity profile:
%
\begin{align}
    1 - \left(A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}}\right) =
    \dfrac{1}{2}\delta
\end{align}
%
Solving for $\Delta\theta$ yields
%
\begin{align}
    \label{eq:hwhm}
    \Delta\theta =
    \cos^{-1} \left[ \dfrac{2 + 3 r' (2 + r')}{2 (1 + r')^3} \right]
    \quad.
\end{align}
%
Differentiation with respect to $r'$ yields
%
\begin{align}
    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r'} =
    \frac{3 r' \left(2 + r'\right)}{\left(1 + r'\right)^4
        \sqrt{\frac{r'^2 \left(2 + r'\right)^2 \left(1 + 2 r'\right)
                \left(3 + 2 r'\right)}{\left(1 + r'\right)^6}}}
    \quad,
\end{align}
%
which is positive definite for all $r' > 0$.

\subsubsection{Behavior at finite degree}
%
The properties of the spot expansion described in the previous section
(in particular its monotonicity and values at the spot center and antipode)
are inherited from the generating function of the Legendre polynomials
(Equation~\ref{eq:gen}). These properties are therefore only rigorously
true when the expansion is taken to spherical
harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
a value of $r'$ in Equation~(\ref{eq:s}) that is too small will lead
to both ringing and a value at the spot center that is in general larger
than $1 - \delta$.

@@

To circumvent these issues, we impose a minimum radius
$r'_{\mathrm{min}}$, whose value is
a function of the degree of the expansion. Since the normalized spot radius
$r$ is defined in the range $[0, 1]$ (Equation~\ref{eq:rprime}), this
minimum value is equal to $c_0$. We compute $c_0 = c_0(l_{\mathrm{max}})$
by numerically finding the
value of $r'$ below which the fractional error in the intensity at the
center of the spot exceeds some threshold, which we take to be 0.01.
Since the width of the spot $\Delta\theta$ approaches $\pi$ as
$r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
maximum radius $r'_{\mathrm{max}}$, which we define to be the radius
corresponding to
$\Delta\theta = \nicefrac{5\pi}{12} = 75^\circ$. This choice is rather
arbitrary, but motivated by the fact that below this value, $\Delta\theta$
scales approximately logarithmically with $r'$ (see Figure~\ref{fig:hwhm}).

\subsection{Probability distribution}
%
Our goal now is to characterize the distribution of $s_{n}$ values as a
function of certain statistics of the spots: specifically, the distribution of
spot radii and intensities. As we will see, if we model these with a Beta and
a log-normal distribution, respectively, we may analytically compute the
first two moments of the distribution of $s_{n}$ conditioned on the parameters
of these two prior distributions.

\subsubsection{Distributions over $r$ and $\delta$}
%
Since we require the normalized spot radius $r$ to be bounded
on the interval $[0, 1]$, we place a Beta prior on this quantity, whose
probability density function (PDF) is
%
\begin{align}
    p \big(r \, \big| \, \alpha, \beta \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    r^{\alpha - 1}
    (1 - r)^{\beta - 1}
    \quad,
\end{align}
%
where $\Gamma$ is the Gamma function and $\alpha$ and $\beta$ are
hyperparameters characterizing the shape of the distribution.
%
And since we require the brightness
%
\begin{align}
    b \equiv 1 - \delta
\end{align}
%
at the center of the spot to
be non-negative, it is convenient to treat $b \in [0, \infty)$ as our
random variable instead of $\delta$. We therefore place a log-normal
prior on $b$, whose PDF is
%
\begin{align}
    \label{eq:lognormal}
    p \big(b \, \big| \, \mu, \nu \big)
     & =
    \frac{1}{b\sqrt{2\pi\nu}}
    \exp\left[
        -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
        \right]
    \quad,
\end{align}
%
where $\mu$ and $\nu$ are the mean and variance of the log-normal, respectively.
The joint distribution over $r$ and $b$ is therefore separable,
by construction:
%
\begin{align}
    p \big( r, b \, \big| \, \alpha, \beta, \mu, \nu \big)
     & =
    p \big(r \, \big| \, \alpha, \beta \big)
    \,
    p \big(b \, \big| \, \mu, \nu \big)
    \quad.
\end{align}
%
In the next two sections, we derive closed form expressions for the first
and second moments of the distribution
%
$p \big( s_{n} \, \big| \, \alpha, \beta, \mu, \nu \big)$.

\subsubsection{First moment}

>>>>

\section{Gaussian Process}
\label{sec:gp}
%

% The vector $\pmb{\theta}$ includes physical properties of the star such
% as its inclination $i$ and rotational period $P$ as well as parameters
% describing the shape of the probability density function (PDF) governing
% the distribution of features on the surface.

Let
$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_K \right)^\top$
denote a vector of $K$ flux measurements at times
$\left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
Conditioned on certain physical properties of the star, $\pmb{\theta}$,
we wish to compute the mean $\pmb{\mu}(\pmb{\theta})$ and
covariance $\pmb{\Sigma}(\pmb{\theta})$
of $\mathbf{f}$, which together fully specify our GP in flux.
%
As with any random variable, the mean and covariance may be computed from
the expectation value of $\mathbf{f}$ and
$\mathbf{f}\,\mathbf{f}^\top$, respectively:
%
\begin{align}
    \label{eq:mean}
    \pmb{\mu}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta} \Big]
    \\
    \label{eq:cov}
    \pmb{\Sigma}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}^2(\pmb{\theta})
\end{align}
%
where the squaring operation in Equation~(\ref{eq:cov}) is
performed element-wise.
%
In \citet{Luger2019} we showed that $\mathbf{f}$ may be computed from a
linear operation on the vector of spherical harmonic coefficients
describing the surface, $\mathbf{y}$:
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \mathbf{A} \, \mathbf{y}
    \quad,
\end{align}
%
where $\mathbf{A}$ is the \starry design matrix, which is implicitly
a function of $\pmb{\theta}$ (as it depends on the stellar inclination
and rotation period, for example).
%
Given Equation~(\ref{eq:fAy}),
we may write the mean and covariance of our flux GP as
%
\begin{align}
    \pmb{\mu}(\pmb{\theta})
     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
    \\
    \pmb{\Sigma}(\pmb{\theta})
     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\Sigma}_{\mathbf{y}} \, \mathbf{A}^\top(\pmb{\theta})
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:mean_y}
    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
    \\
    \label{eq:cov_y}
    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})
     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}_{\mathbf{y}}^2(\pmb{\theta})
\end{align}
%
are the mean and covariance of the GP in the spherical harmonics basis.
The bulk of the math in this paper is devoted to computing
the expectations in the expressions above, which
are given by the integrals
%
\begin{align}
    \label{eq:exp_y}
    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
    \\
    \label{eq:exp_yy}
    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
    \quad,
\end{align}
%
where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
distribution of features on the surface  (i.e., the size and location of star spots)
and $p(\mathbf{x} \, \big| \, \pmb{\theta})$ is its probability density
function (PDF).

In the sections that follow we will show that for suitable choices of $\mathbf{y}(\mathbf{x})$

@@

    \mathbf{x} = \left( \lambda \,\, \phi \,\, \xi \,\, r \right)^\top

@@

    \xi
    \mathbf{s}(r)

@@

where $\lambda$ is the longitude of a spot, $\phi$ is its latitude, $\xi$
is its contrast, and $r$ is its radius. The vector function $\mathbf{s}(r)$
returns the spherical harmonic expansion of a negative unit brightness
circular spot at $\lambda = \phi = 0$,

@@

$\mathbf{x}$ describing the size, brightness, and position of a spot to the

@@

    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,

@@

    \pmb{\theta} = \left(
    \pmb{\theta}_{\lambda} \, \,
    \pmb{\theta}_{\phi} \, \,
    \pmb{\theta}_{\xi} \, \,
    \pmb{\theta}_{r} \right)^\top

@@

    \mathbf{e_4}(\pmb{\theta})

@@

    \mathbf{E_4}(\pmb{\theta})

@@

    \label{eq:e1}
    \mathbf{e_1}(\pmb{\theta}_r)
     & \equiv
    \mathbf{s}(r) \,

@@

    \mathrm{d}r
    \label{eq:e2}
    \mathbf{e_2}(\pmb{\theta}_b, \mathbf{e_1})
     & \equiv
    \xi \,
    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
    \mathrm{d}\xi

@@

    \label{eq:e3}
    \mathbf{e_3}(\pmb{\theta}_\phi, \mathbf{e_2})
     & \equiv

@@

    \label{eq:e4}
    \mathbf{e_4}(\pmb{\theta}_\lambda, \mathbf{e_3})
     & \equiv

@@

    \mathbf{e_3} \,

@@

    \label{eq:E1}
    \mathbf{E_1}(\pmb{\theta}_r)
     & \equiv
    \mathbf{s}(r) \, \mathbf{s}^\top(r) \,

@@

    \mathrm{d}r

@@

    \label{eq:E2}
    \mathbf{E_2}(\pmb{\theta}_b, \mathbf{E_1})
     & \equiv
    \xi^2 \,
    p(\xi \, \big| \, \pmb{\theta}_b)
    \mathrm{d}\xi
    \label{eq:E3}
    \mathbf{E_3}(\pmb{\theta}_\phi, \mathbf{E_2})
     & \equiv

@@

    \label{eq:E4}
    \mathbf{E_4}(\pmb{\theta}_\lambda, \mathbf{E_3})
     & \equiv

@@

    \mathbf{E_3} \,
    \mathbf{E_3}^\top \,

@@

We devote the next four sections to the computation of these eight

@@

%
\section{The Size Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
\label{sec:size}
%
\subsection{Functional form}
\label{sec:size-function}
%
We adopt the following expression for the $n^{\mathrm{th}}$ term in $\mathbf{s}$
in the expansion of a dark, unit contrast circular spot at
$\lambda = \phi = 0$:
%
\begin{align}
    \label{eq:s}
    s_{n}(r) =
    \begin{cases}
        \dfrac{r' \delta_{l0}}{2}
        -\dfrac{r' \left( 2 + r' \right)}
        {2 \sqrt{2l + 1} (1 + r')^{l + 1}}
         & m = 0
        \\
        0
         & \mathrm{otherwise}
    \end{cases}
\end{align}
%
where
$l = l(n), m = m(n)$ (Equation~\ref{eq:lm}) and
$\delta$ is the Kronecker delta. The parameter $r'$ is defined as
%
\begin{align}
    \label{eq:rprime}
    r' \equiv c_0 + c_1 r
    \quad,
\end{align}
%
where $c_0, c_1 > 0$ are scaling constants (defined in more detail below)
and $r \in [0, 1]$ is the normalized spot radius.
The expression in Equation~(\ref{eq:s}) is convenient because it satisfies
five important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity is azimuthally symmetric about the spot center
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $-1$
    \item The surface intensity at the antipode of the spot center is zero
    \item The size of the spot increases monotonically with $r$
\end{enumerate}
%
These properties may be demonstrated by considering the
expression for the surface intensity at a given polar angle $\theta$ and
azimuth $\varphi$:
%
\begin{align}
    \begin{split}
        I(\theta, \varphi)
        & =
        \sum\limits_{n=0}^\infty
        s_{n} Y_{n}(\theta, \varphi) \\
        & =
        \sum\limits_{n=0}^\infty
        s_{n} \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
    \end{split}
\end{align}
%
where $Y_n$ is the spherical harmonic of degree $l(n)$ and order $m(n)$
(see Equation~\ref{eq:lm}),
$P_l$ is the Legendre polynomial of degree $l$,
and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Note that the expression above is independent of the azimuth $\varphi$,
a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
expansion is therefore azimuthally symmetric about the spot center, as
stated in \textbf{1}.
%
Combining the above expression with Equation~(\ref{eq:s}) and
rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    I(\theta) =
    \dfrac{r'}{2}
    -
    \dfrac{r' \left( 2 + r' \right)}{2 (1 + r')}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + r'}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity as a function of polar angle
in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
    %
    \\
    \intertext{where}
    %
    \begin{split}
        A & = \dfrac{r'}{2}                      \\
        B & = r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
        C & = \dfrac{1 + (1 + r')^2}{2 + 2 r'}
    \end{split}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{2}. The value at the minimum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
        & = -1
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{3}, and the value at the maximum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
        & = 0
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{4}.
Finally, to show \textbf{5}, let us compute the half width at half minimum
$\Delta\theta$ of the intensity profile:
%
\begin{align}
    A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}} =
    -\dfrac{1}{2}
\end{align}
%
Solving for $\Delta\theta$ yields
%
\begin{align}
    \label{eq:hwhm}
    \Delta\theta =
    \cos^{-1} \left[ \dfrac{2 + 3 r' (2 + r')}{2 (1 + r')^3} \right]
    \quad.
\end{align}
%
Differentiation with respect to $r'$ yields
%
\begin{align}
    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r'} =
    \frac{3 r' \left(2 + r'\right)}{\left(1 + r'\right)^4
        \sqrt{\frac{r'^2 \left(2 + r'\right)^2 \left(1 + 2 r'\right)
                \left(3 + 2 r'\right)}{\left(1 + r'\right)^6}}}
    \quad,
\end{align}
%
which is positive definite for all $r' > 0$.

@@

It is important to note that the
properties of the spot expansion described above
(in particular its monotonicity and values at the spot center and antipode)
are inherited from the generating function of the Legendre polynomials
(Equation~\ref{eq:gen}). These properties are therefore only rigorously
true when the expansion is taken to spherical
harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
a value of $r'$ in Equation~(\ref{eq:s}) that is too small will lead
to both ringing and a value at the spot center that is in general larger
than $-1$.

To circumvent these issues, we impose a minimum radius
$r'_{\mathrm{min}}$, whose value is
a function of the degree of the expansion. Since the normalized spot radius
$r$ is defined in the range $[0, 1]$ (Equation~\ref{eq:rprime}), this
minimum value is equal to $c_0$. We compute $c_0 = c_0(l_{\mathrm{max}})$
by numerically finding the
value of $r'$ below which the fractional error in the intensity at the
center of the spot exceeds some threshold, which we take to be 0.01.
Since the half width of the spot $\Delta\theta$ approaches $\nicefrac{\pi}{2}$ as
$r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
maximum radius $r'_{\mathrm{max}}$, which we define to be the radius
corresponding to
$\Delta\theta = \nicefrac{5\pi}{12} = 75^\circ$. This choice is rather
arbitrary, but motivated by the fact that below this value, $\Delta\theta$
scales approximately logarithmically with $r'$.

The left panel of Figure~\ref{fig:hwhm} shows this scaling. The minimum
and maximum values of $r'$ for $l_{\mathrm{max}} = 20$ are indicated
for reference; these correspond to $r = 0$ and $r = 1$, respectively
(top axis). The right panel of the figure shows the minimum spot half width
$\Delta\theta$ as a function of the spherical harmonic degree of the
expansion for different error thresholds: ten percent (dashed), one
percent (solid), and one part per thousand (dotted).
For $l_{\mathrm{max}} = 20$, the minimum spot half width is about 17$^\circ$.
To model spots smaller than this, one must go to larger $l_{\mathrm{max}}$;
note, however, how slowly $\Delta\theta$ decreases above
$l_{\mathrm{max}} = 20$.

The intensity as a function of polar angle away from the spot center
is shown in Figure~\ref{fig:spot_expansion} for a range of $r$ values.
Three spots with $r < 0$ are also shown; note the presence of
ringing and the fact that the intensity at the spot center is significantly
larger than $-1$.

\subsection{PDF}
\label{sec:size-pdf}
%
Our goal now is to characterize the distribution of $\mathbf{s}$ as a
function of parameters $\pmb{\theta}_r$ describing the distribution of
spot radii. Since $r \in [0, 1]$, we choose to model $r$ as a random
Beta-distributed variable; as we will see, this choice will allow us to
analytically compute the first two moments of the distribution of
$\mathbf{s}$ conditioned on $\pmb{\theta}_r$.

%
The Beta distribution in $r$ has hyperparameters
%
\begin{align}
    \pmb{\theta}_r = \left(
    \alpha \, \, \, \,
    \beta \right)^\top
\end{align}
%
and PDF given by
%
\begin{align}
    \label{eq:pdf_r}
    p \big(r \, \big| \, \pmb{\theta}_r \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    r^{\alpha - 1}
    (1 - r)^{\beta - 1}
    \quad,
\end{align}
%
with support in $0 \leq r \leq 1$,
where $\Gamma$ is the Gamma function.
